---
title: "Identifying Undervalued Advertising Keywords"
author: fastzhong
highlighter: highlight.js
output:
  html_document:
    keep_md: yes
job: null
knit: slidify::knit2slides
mode: standalone
hitheme: tomorrow
subtitle:
framework: io2012
widgets: []
---

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(wordcloud)

if(!file.exists('data')) {
  dir.create('data' )
}
if( !file.exists('data/US-market' )) {
  if( !file.exists('data/bids-dataset-1.zip')) {
    download.file('https://open-advertising-dataset.googlecode.com/files/bids-dataset-1.zip',
                  destfile='data/bids-dataset-1.zip', method='curl')
  }
  unzip('data/bids-dataset-1.zip', exdir='data')
}

df <- read.csv('data/US-market/keyword_stats_20120531_0148831.csv', na.strings='-')
colnames(df) <- c('keyword', 'global_monthly_searches', 'est_avg_cpc', 'est_ad_pos', 
                  'est_daily_clicks', 'est_daily_cost', 'local_monthly_searches')
df$est_daily_cost <- as.numeric(gsub( ',', '', df$est_daily_cost))
df <- filter(df, !is.na( global_monthly_searches), !is.na(local_monthly_searches))
```

## Summary

Search engines allow advertisers to bid on keywords. Some are more expensive and some 
are submitted more often. What are the most cost-effective keywords to bid?

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.cap='Common Keywords'}
wordcloud(unlist(strsplit(as.character(df$keyword), ' ')))
```

--- .class #id

## Advertising Bidding Data
Notice the skewed distribution of continuous variables.
```{r}
summary(df[ ,2:length(df)])
```

--- .class #id 

## Distributions Illustrated

```{r echo=FALSE}
par(mfrow=c(3, 2))
hist(log(df$est_avg_cpc + 1 ), main='Cost per Click')
hist(df$est_ad_pos, main='Estimated Ad Position')
hist(log(df$est_daily_clicks + 1), main='Estimated Daily Clicks')
hist(log(df$est_daily_cost + 1 ), main='Estimated Daily Cost')
hist(log(df$local_monthly_searches + 1),
     main='Local Monthly Searches')
```

--- .class #id 

## Conclusion

Because the search frequency does not follow the same distribution of the cost 
there may be undervalued search keywords. This application will help identify 
those keywords.

https://fastzhong.shinyapps.io/developing-data-products/
